---
title: "Context - Problem Understanding"
author: "Simon Hulme"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
diabetes_data <- read_rds("00_Data/processed/for_analysis/diabetes_for_analysis.rds")
```

# Determine Business Objectives

## Background

```{r}
diabetes_data %>% 
    count(readmitted) %>% 
    mutate(percentage = round(n / sum(n), 3) * 100)
```

-   Patients with diabetes more likely to be readmitted early

    -   known risk factors

        -   socio-demographic, co-morbidities, insulin therapy, length
            of stay, and previous admissions.

-   In this sample 8.2% of patients with diabetes were readmitted within
    30 days of discharge

    -   lower rate than expected suggesting either data incomplete or
        issues with existing evidence

-   Early readmissions are costly

    -   total NHS cost

    -   cost per organisation

    -   'cost' to patients - not necessarily financial

## Business Objectives

-   Reduce the costs associated with early readmission of patients with
    diabetes

-   Select appropriate intervention to reduce readmission

    -   low cost applied to many patients

    -   high cost targeted at high-risk patients

## Business Success Criteria

-   Minimise overall cost

    -   overall cost = cost attributed to all early readmissions + cost
        of intervention

# Assess situation

## Inventory of resources

-   personnel: Me (Ex GP and Data Science Consultant)

-   data: secondary care data from US 1999-2008 (UCI Machine Learning
    Repository)

-   computing: iMac

-   software: R (packages: tidyverse, h2o, recipes)

## Requirements, assumptions, and constraints

-   requirements

    -   schedule of completion,

    -   comprehensibility

    -   quality of results

    -   security

    -   legal issues

-   assumptions

    -   patients are adults discharged to community settings (excluding
        hospices)

        -   exclude patients discharged to other hospitals or hospital
            departments

        -   exclude patients who die

        -   exclude patients discharged for end-of-life care

    -   these may be assumptions about the data that can be verified
        during data mining,

    -   may also include non-verifiable assumptions about the business
        related to the project.

    -   It is particularly important to list the latter if it will
        affect the validity of the results.

-   constraints

    -   availability of resources

    -   technological constraints such as the size of dataset that it is
        practical to use for modeling.

Risks and contingencies

-   risks or events that might delay the project or cause it to fail.

-   corresponding contingency plans, what action will be taken if these
    risks or events take place.

## 

Terminology

-   relevant business terminology

-   data mining terminology

Costs and benefits

-   compares the costs of the project with the potential benefits to the
    business if it is successful.

    -   be as specific as possible. For example, use monetary measures
        in a commercial situation.

# Determine data mining goals

## Data mining goals

-   predict which patients with diabetes admitted to hospital are at
    increased risk of early hospital readmission

-   understand drivers for readmission to aid development of an
    appropriate intervention

-   stratify risk using predicted probability

## Data mining success criteria

-   tbd

-   *Define the criteria for a successful outcome to the project in
    technical terms—for example, a certain level of predictive accuracy
    or a propensity-to-purchase profile with a given degree of “lift.”
    As with business success criteria, it may be necessary to describe
    these in subjective terms, in which case the person or persons
    making the subjective judgement should be identified.*

# Produce project plan

## Project plan

-   tbd

*Describe the intended plan for achieving the data mining goals and
thereby achieving the business goals. The plan should specify the steps
to be performed during the rest of the project, including the initial
selection of tools and techniques.*

*List the stages to be executed in the project, together with their
duration, resources required, inputs, outputs, and dependencies. Where
possible, make explicit the large-scale iterations in the data mining
process—for example, repetitions of the modeling and evaluation phases.*

*As part of the project plan, it is also important to analyze
dependencies between time schedule and risks. Mark results of these
analyses explicitly in the project plan, ideally with actions and
recommendations if the risks are manifested. Note: the project plan
contains detailed plans for each phase. Decide at this point which
evaluation strategy will be used in the evaluation phase.*

*The project plan is a dynamic document in the sense that at the end of
each phase, a review of progress and achievements is necessary and a
corresponding update of the project plan is recommended. Specific review
points for these updates are part of the project plan.*

## 

Initial assessment of tools and techniques

-   tbd

*At the end of the first phase, an initial assessment of tools and
techniques should be performed. Here, for example, you select a data
mining tool that supports various methods for different stages of the
process. It is important to assess tools and techniques early in the
process since the selection of tools and techniques may influence the
entire project.*
